---
title: "Statistical assignment 1"
author: "Andrew Bosbery"
date: "02/02/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open data (10 points)

In this assignment you will work with the individual level data from wave 8 of the Understanding Society survey. First, you need to open the data set. Please complete the code below.

```{r}
library(tidyverse)
Data <- read_tsv("C:/Users/Andrew/Documents/University/Year 2/Term 2/Data Analysis 3/Project/data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
```

Now you have got your data frame stored as Data. 

## Select variables (10 points)

The data for Wave 8 of the Understanding Society were collected in 2016-18. Among other things, people were asked the following question: "Should the United Kingdom remain a member of the European Union or leave the European Union?" In this assignment, we will explore how answers to this question depend on sex and age.

First, you need to select the variables for the analysis. You want to keep the following variables: cross-wave individual identifier (*pidp*), support for the UK remaining or leaving the EU (*h_eumem*), sex (*h_sex_dv*), age (*h_age_dv*), and sample origin (*h_memorig*).

Complete the code below to select those variables from the data frame and save the result.

```{r}
Brexit <- Data %>%
        select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)
```

## Filter observations (10 points)

To make nationally representative estimates from the Understanding Society data we would need to use weight coefficients. There are many different types of weight coefficients that can be used depending on the question and the level of analysis (see the User Guide, pp. 65-71). We will not do this in this assignment. However, what we want to do is to keep data from the original Understanding Society sample only (ukhls gb 2009-10), dropping  data for Northern Ireland, the BHPS cohort members and ethnic minority boost samples. This will make data closer to be representative for Great Britain. You need to choose the observations where *h_memorig* has the value of 1.

```{r}
Data <- Data %>%
        filter(Data, h_memorig == 1)
```

## Recode data (20 points)

Let us tabulate the variables for EU support, sex, and age.

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)
```

You will see that all these variables are numeric. You can learn what the numeric codes mean by checking the codebook here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/datafile/indresp/wave/8 .

We want to do the following:

1) Recode the variable for EU support as binary (1 for Remain, 0 for Leave), coding all types of missing values (including refusals and "don't know") as NA.
2) Recode sex into a character vector with the values "male" or "female".
3) Recode age into a variable with the following categories: 16 to 25, 26 to 40, 41 to 55, 56 to 70, over 70.

In all cases, we want to create new variables.

```{r}
Data <- Data %>%
        mutate(EU = ifelse(h_eumem == 1, 1,
                           ifelse(h_eumem == 2, 0, NA))     
      
          ) %>%
        mutate(sex = case_when(
          h_sex_dv == 1 ~ "Male",
          h_sex_dv == 2 ~ "Female")
          
           ) %>%
        mutate(agegr = case_when(
          between(h_age_dv, 16, 25) ~ "16 to 25",
          between(h_age_dv, 26, 40) ~ "26 to 40",
          between(h_age_dv, 41, 55) ~ "41 to 55",
          between(h_age_dv, 56, 70 ) ~ "56 to 70",
          h_age_dv > 70 ~ "Over 70"
        ))
        

```

## Summarise data (20 points)

Let us **dplyr** to calculate how many people in the sample supported Remain and Leave, both as absolute numbers and percentages.

```{r}
Data %>%
 count(EU, name = "Number of Votes")   %>%
 mutate(Remain_Perc = (length(Data$EU[Data$EU == 1]) - length(Data$EU[is.na(Data$EU)])) / (length(Data$EU) - length(Data$EU[is.na(Data$EU)])) * 100)  %>%
 mutate(Leave_Perc = (length(Data$EU[Data$EU == 0]) - length(Data$EU[is.na(Data$EU)])) / (length(Data$EU) - length(Data$EU[is.na(Data$EU)])) * 100)

  
    #This worked but was highly inefficient    
```

Write a couple of sentences with the interpretation of this result. How this compares with the result of the 2016 referendum? Why?

This suggests that the majority of individuals in this survey wanted to remain in the EU whereas in the 2016 referendum a narrow majority voted to leave (58.5% Remain in the survey compared to 48% Remnain in the Referendum). Some important reasons for this may be that this survey included 16 and 17 year olds. These individuals could not vote and if they heavily favoured remaining in the EU than this could skew the result in the survey. Furthermore the survey did not attempt to predict turnout, if the more remain voters did not turn out to vote than leave voters than this would also affect the results of the survey. This would be especially true if many remain supporters were younger people who tend to have a lower turnout than older voters who tender to vote to leave.

## Summarise data by sex and age (30 points)

Now let us look at the support for Leave and Remain by sex and age. Use your newly created variables.


```{r}
Data %>%
  filter(sex == "Male")  %>%
  #mutate(remain_perc_male = mean(EU, na.rm = TRUE)) * 100  %>%
  group_by(agegr)   %>%
  summarise(remain_perc_male = mean(EU, na.rm = TRUE) * 100)
  
Data %>%
  filter(sex == "Female")  %>%
  group_by(agegr)  %>%
  summarise(remain_perc_female = mean(EU, na.rm = TRUE) * 100)






#summarise(remain_perc_male = mean(EU, na.rm = TRUE) * 100) 
#tapply(Data$EU, Data$sex, mean, na.rm = TRUE)
#tapply(Data$EU, Data$agegr, mean, na.rm = TRUE)
        
```


Write a couple of sentences interpreting your results.
Overall both men and women supported remaing in the EU. Althogh 60% of women suprrorted remaining as opposed to 55.6% of men. This suggests that there may be a link between gender and support of membership of the EU. This seems to be true as at every age range a higher percentage of women supported remain over leave. Also The younger the respondant the more likely they were to support membership of the EU. In fact in only the over 70s category did more people oppose EU membership than support it. This evidence backs up the earlier stated interpretation that due to 16 and 17 year olds being unable to vote and that turnout is less amongst younger voters that the one reason why the UK voted to leave the EU was a lack of voters turning out in younger demographics.